A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
added ID3 initial description.

This tool was used to apply the ID3 \cite{id3}\cite{1} algorithm, mainly because
of its support for numerical attributes (not seen in Weka), wich enabled a
better comparison of approaches.

@@

@@

\begin{table}
\begin{center}
\begin{tabular}{ | l | l | l | l | l |}
    \hline
    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    Unordered & Laplacian & 13.3 & 15.2 & 3min29sec \\ \hline
    Unordered & Naive & - & - & - \\ \hline
    Ordered & Laplacian & 0 & 17.6 & 1min45sec \\ \hline
    Ordered & Naive & 0 & 19.9 & 12min21sec \\ \hline
    Ordered & Information Gain & 18.9 & 18.9 & 5sec \\ \hline
    Ordered & Modified Entropy & 18.9 & 19.1 & 20sec \\ \hline
\end{tabular}
\caption{Results for CN2.}
\label{tbl:results_cn2}
\end{center}
\end{table}

>>>>

This tool was used to apply the ID3 \cite{id3}\cite{1} algorithm, mainly because
of its support for numerical attributes (not seen in Weka), which enabled a
better comparison of approaches.

The data pre-processing conducted consisted on a missing value replacement by 
the average on numerical values and by the mode on the nominal/ordinal values. 
An absolute discretization was applied to the numerical values for testing
purposes of the default ID3 implementation.

@@

\subsection{ID3}

ID3 in RapidMiner is implemented in two different flavours, one that supports
numerical values, one that doesn't. For the one that doesn't support numerical
values, an absolute discretization was conducted. Table \ref{tbl:results_id3}
summarizes the results with the ID3 algorithm.

\begin{table}
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Type} & \textbf{Criterion} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    ID3 & Gain Ratio & 0 & 0 & 0 \\ \hline
    ID3 & Information Gain & 0 & 0 & 0 \\ \hline
    ID3 & Gini Index & 0 & 0 & 0 \\ \hline
    ID3 & Accuracy & 0 & 0 & 0 \\ \hline
    ID3 Numerical & Gain Ratio & 0 & 0 & 0 \\ \hline
    ID3 Numerical & Information Gain & 0 & 0 & 0 \\ \hline
    ID3 Numerical & Gini Index & 0 & 0 & 0 \\ \hline
    ID3 Numerical & Accuracy & 0 & 0 & 0 \\ \hline
  \end{tabular}
  \caption{Results for ID3.}
  \label{tbl:results_id3}
  \end{center}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{ | l | l | l | l | l |}
    \hline
    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    Unordered & Laplacian & 13.3 & 15.2 & 3m29s \\ \hline
    Unordered & Naive & - & - & - \\ \hline
    Ordered & Laplacian & 0 & 17.6 & 1m45s \\ \hline
    Ordered & Naive & 0 & 19.9 & 12m21s \\ \hline
    Ordered & Information Gain & 18.9 & 18.9 & 0m5s \\ \hline
    Ordered & Modified Entropy & 18.9 & 19.1 & 0m20s \\ \hline
\end{tabular}
\caption{Results for CN2.}
\label{tbl:results_cn2}
\end{center}
\end{table}
